# Reproducible Research: Peer Assessment 1

## Loading and preprocessing the data

1. Loading
    ```{r unzipit}
    # Unzip to temp file, because there's no need to clog the working directory
    here <- tempdir()
    unzip('repdata-data-activity.zip',exdir=here)

    # Read into data frame and see what's in it
    activity <- read.csv(paste(here,'activity.csv',sep='/'))
    ```
2. Preprocessing
    ```{r preprocessit}
    head(activity)
    str(activity)

    # Make date strings into proper dates
    activity$date <- as.Date(activity$date)
    str(activity)
    ```

## What is mean total number of steps taken per day?

1. Histogram of the number of steps:

    ```{r histogram}
    hist(activity$steps)
    meanSteps <- mean(activity$steps,na.rm=TRUE)
    medianSteps <- median(activity$steps,na.rm=TRUE)
    zeroSteps <- sum(activity$steps==0,na.rm=TRUE)
    ```
    
2. Mean and median:

Ignoring missing values, the mean number of steps taken per day is `r round(meanSteps,2)` and the median is `r medianSteps`. Is `r medianSteps` right? There are `r prettyNum(nrow(activity),big.mark=",")` observations. Of these, `r prettyNum(zeroSteps,big.mark=",")` record 0 steps. That's `r round(100*zeroSteps/nrow(activity),2)`%, so it looks right.

## What is the average daily activity pattern?

1. Plot of average steps by 5-minute intervals:

    ```{r stepplot}
    ms <- aggregate(activity, by=list(activity$interval), FUN=mean, na.rm=TRUE)
    ms <- subset(ms,select=c(steps,interval))
    plot(ms$steps~ms$interval,type='l',main='Mean steps by 5-minute interval',
    ylab="Steps",xlab="Interval",las=1)
    ```
    
2. The interval containing the maximum number of steps is `r ms$interval[ms$steps==max(ms$steps)]`.

## Imputing missing values

1. The total number of rows with missing values in the data set is `r nrow(activity[is.na(activity),])`. There are three columns, and in theory any of them could have missing values. The total number of cases where only the number of steps is missing is `r sum(is.na(activity$steps))`. I can check that there are no missing values in any of the other two variables: date (`r sum(is.na(activity$date))` missing values) or interval (`r sum(is.na(activity$interval))` missing values).

2. Filling in missing values: I will just use means by interval that I already computed. That's my strategy.

3. Creating a new data set with filled in values:

    ```{r imputemissing}
    # Bring in means by 5-minute interval
    foo <- merge(activity,ms,by=c('interval'))
    
    # Swap means for missings
    foo$steps <- foo$steps.x
    foo$steps[is.na(foo$steps.x)] <- foo$steps.y[is.na(foo$steps.x)]
    
    # Tidy up:  
    foo <- subset(foo,select=-c(steps.x,steps.y)) # drop unneeded columns
    foo <- foo[order(foo$date),]                  # sort by date
    foo <- foo[,c(3,2,1)]                         # restore original column order
    
    # New mean and median number of steps
    meanSteps <- mean(foo$steps)
    medianSteps <- median(foo$steps)
    zeroSteps <- sum(foo$steps==0)
    ```

4. With imputed missing values, the mean number of steps taken per day is `r round(meanSteps,2)` and the median is `r round(medianSteps,2)`. Is `r medianSteps` right? There are `r prettyNum(nrow(foo),big.mark=",")` observations. Of these, `r prettyNum(zeroSteps,big.mark=",")` record 0 steps. That's `r round(100*zeroSteps/nrow(foo),2)`%, so it looks right. 

Imputing missing values with this strategy has no effect on the mean (effectively, the mean over non-missings is filled in for missings). It also happens to have no effect at all on the median, because zero values amount to more than half of all observations both before and after filling in the missing values. The new histogram is below:

```{r newhistogram}
hist(foo$steps)
```

## Are there differences in activity patterns between weekdays and weekends?

1. Tag weekend data:

    ```{r weekdays}
    # Tag weekends and check
    foo$weekday <- weekdays(as.Date(foo$date))
    foo$weekend <- foo$weekday %in% c('Sunday','Saturday')
    table(foo$weekday,foo$weekend)
    ```
2. Plot it:
    
    ```{r plotweekends}
    # Panel plot
    require(ggplot2)
    foo$weekend <- factor(foo$weekend,labels=c('weekday','weekend'))
    p <- ggplot(foo, aes(interval,steps)) + 
        geom_line() + facet_grid(weekend~.) +
        ggtitle('Walking patterns')
    p
    ```

It appears that on weekends people wake up later, because the average number of steps picks up later in the day than it does on weekdays.
