Peer Assessment 1
========================================================

## Loading and preprocessing the data

1. Loading
    ```{r unzipit}
    # Unzip to temp file, because there's no need to clog the working directory
    here <- tempdir()
    unzip('repdata-data-activity.zip',exdir=here)

    # Read into data frame and see what's in it
    activity <- read.csv(paste(here,'activity.csv',sep='/'))
    ```
2. Preprocessing
    ```{r preprocessit}
    head(activity)
    str(activity)

    # Make date strings into proper dates
    activity$date <- as.Date(activity$date)
    str(activity)
    ```

## What is mean total number of steps taken per day?

1. Histogram of the number of steps:
    ```{r histogram, echo=FALSE}
    hist(activity$steps)
    meanSteps <- mean(activity$steps,na.rm=TRUE)
    medianSteps <- median(activity$steps,na.rm=TRUE)
    zeroSteps <- sum(activity$steps==0,na.rm=TRUE)
    ```
2. Mean and median:

Ignoring missing values, the mean number of steps taken per day is `r round(meanSteps,2)` and the median is `r medianSteps`. Is `r medianSteps` right? There are `r nrow(activity)` observations. Of these, `r zeroSteps` record 0 steps. That's `r round(100*zeroSteps/nrow(activity),2)`%, so it looks right.

## What is the average daily activity pattern?
